---
title: "Lab 03 - Nobel laureates"
author: "John T. BennettA"
date: "Feb. 10, 2022"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1

There are 935 observations in the dataset. The number of variables in the dataset is 26. Each row represents one Nobel laureate. 


```{r viewdata}
View(nobel)
```

### Exercise 2

2.1 Create dataframe
    Create a new data frame called nobel_living that filters for:
    -laureates for whom country is available
    -laureates who are people as opposed to organizations (organizations are denoted with "org" as their gender)
    -laureates who are still alive (their died_date is NA)

     Confirm that once you have filtered for these characteristics you are left with a data frame with 228 observations, once again using inline code.
     
     **JTB Note: In order to see the actual 228x26 tibble, I've entered the code without the dataframe command (myDataframe <-) immediately below. Then, below that, I've entered the code with the dataframe command. 

```{r newdataframe1}
nobel %>%
  filter(!is.na(country),
  gender != "org",
  is.na(died_date))
```

```{r newdataframe2}
nobel_living <- nobel %>%
  filter(!is.na(country),
  gender != "org",
  is.na(died_date))
```

2.2 Mutate- 

```{r NobelMutateEx2}
nobel_living <- nobel_living %>%
  mutate(
    country_us = if_else(country == "USA", "USA", "Other")
  )
```

2.3 Mutate- Physics, Medicine, Chemistry, and Economics.

```{r NobelMutate_by_field}
nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
```


### Exercise 3

This visualization shows whether the laureate was in the US when they won the Nobel prize. This shows that a large majority of laureates were indeed in the US when they won their prizes. To know whether the article's main claims are warranted, we would need to know more about the place of birth of the Nobel winners. The "born_country" variable would be better suited for this question. 
If, theoretically, all of the Nobel winners who were in in the US when they were awarded were in fact born in the US, then that would not indicate anything about the influence of immigration. The "born_country" variable is therefore germane to the question. 
    

```{r FacetedBarPlot3}
ggplot(nobel_living_science, aes(x = country_us)) +
 geom_bar() +
 labs(
   x = "Location of Laureate at Time of Award",
   y = "Frequency (Number of Awards)",
   title = "Faceted bar plot"
 ) +
 facet_wrap(~ category)+
coord_flip()
```

 
### Exercise 4
How many of the winners are born in the US? 147 of the winners were born in the US. 

```{r born_country_us}
nobel_living_science <- nobel_living_science %>%
  mutate(
    born_country_us = if_else(country_us == "USA", "USA", "Other")
  )
nobel_living_science %>% count(born_country_us, sort = TRUE)
```


### Exercise 5
Issue: This visualization is not showing any distinctions within the bars.


```{r FacetedBarPlot5}
ggplot(nobel_living_science, 
mapping = aes(x = country_us, fill = born_country_us)) +
 geom_bar() +
 labs(
   x = "Location of Laureate at Time of Award",
   y = "Frequency (Number of Awards)",
   title = "Faceted bar plot",
 ) +
 facet_wrap(~ category)+
coord_flip()
```





### Exercise 6
Issue: In the code below, the commands cause a "0 rows" response. This should not be happening because my "born_country" value of "Other" should not have any problems.

In a single pipeline, filter for laureates who won their prize in the US, but were born outside of the US, and then create a frequency table (with the count() function) for their birth country (born_country) and arrange the resulting data frame in descending order of number of observations for each country. Which country is the most common?


```{r count_and_desc_order}
nobel_living_science %>%
filter(country == "USA", born_country == "Other") %>%
count(born_country) %>%
arrange(desc(born_country))
```
 

